A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution

### 2019-08-02 paper/comments/reply-letter.tex
Most of the reviewer's comments have been addressed.
Revisions done all together.

\noindent We thank all reviewers for their very helpful comments and suggestions. We have addressed all the raised issues. Details are provided below, where we list all reviews together with a detailed response to each suggestion by the reviewers.

@@

\tcomment{General comment: It should be also interesting to compare the crossing resolution between the "non-grid" and the "grid" versions of your algorithm}

\response{\textcolor{red}{Which version of the algorithm do we use in the experiments?}}
\end{itemize}

@@

\response{We preferred not to pose such exact questions, because we simply cannot provide answers to any of them. The reason is that an experimental evaluation can only give some indications. Instead, we now explicitly mention which are the parameters that are interested~in.}

@@

\response{As mentioned in the previous reply, our experimental evaluation aims in giving some indications. The mean is a good metric in this direction, even though it is sensitive to outliers. \textcolor{red}{For completeness, we have added a diagram illustrating the standard deviation as suggested by the reviewer.}}

@@

\response{We are always comparing mean numbers. This is explicitly stated at the beginning of Section 3. Here, we intended to distinguish between some graphs with small number of vertices ($<20$) and the remaining ones of the experiment. In the revised version, we are more precise in our description by explicitly mentioning this. We have revised accordingly.}

>>>>

\noindent We thank all reviewers for their very helpful comments and suggestions. We have addressed all the raised issues. Details are provided below, where we list all reviews together with a detailed response to each suggestion by the reviewers. For convenience, we also attach a diff PDF highlighting the differences between the submitted version and the revision.

@@

\rcomment{General comment: It should be also interesting to compare the crossing resolution between the "non-grid" and the "grid" versions of your algorithm}

\response{In the experiments, since we are comparing our approach with spring embedding algorithms, which do not produce grid drawings, we indeed report results only for the non-grid variant of our algorithm. Only if there exist strong constraints on the size of the underlying grid, one can observe differences between the grid and non-grid variants  in terms of the crossings resolution of the produced drawings, which are again not big (less than 2 degrees). Hence, plotting these differences results in two almost overlapping curves.}
\end{itemize}

@@

\response{We preferred not to pose such exact questions, because we simply cannot provide answers to any of them. The reason is that an experimental evaluation can only give some indications. Instead, we now explicitly mention which are the parameters that are interested~in (indicated by P.1-P.5 in the revision).}

@@

\response{As mentioned in the previous reply, our experimental evaluation aims in giving some indications. The mean is a good metric in this direction, even though it is sensitive to outliers as the reviewer points out. For completeness, we added in our diagrams the standard deviation as suggested by the reviewer. Note that we had to use colors in our figures, as otherwise the amount of information to be visualized in grayscale made the figures hard to read.}

@@

\response{We are always comparing mean numbers. This is explicitly stated at the beginning of Section 3. Here, we intended to distinguish between some graphs with small number of vertices ($<20$) and the remaining ones of the experiment. In the revised version, we are more precise in our description by explicitly mentioning this. }

### 2019-08-02 paper/journal.tex
agreed drawing style.
Revisions done all together.

In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing. How would one increase the crossing resolution of a given drawing? First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase. Of course, we cannot consider all possible positions for the vertex to be moved. Instead, we consider a small set of randomly generated ones. If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position. Note at that in general the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems; see, e.g.,~\cite{DBLP:conf/sofsem/BlasiusRR17,DBLP:conf/dac/FiducciaM82,6771089,DBLP:conf/alenex/RadermacherRRW18,DBLP:conf/wsdm/UganderB13}.

@@

% ==================================================================
%\input{cr-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=1]{figures/rome}}
	\hfil
	\subfloat[\label{fig:cr-res-2}{Edge-length ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=2]{figures/rome}}

	\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=3]{figures/rome}}
	\hfil
	\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=4]{figures/rome}}
	\caption{Experimental results on the crossing resolution for the Rome graphs.}
	\label{fig:cr-res}
\end{figure*}

% ==================================================================
%\input{to-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=5]{figures/rome}}
	\hfil
	\subfloat[\label{fig:to-res-2}{Edge-length ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=6]{figures/rome}}

	\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=7]{figures/rome}}
	\hfil
	\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=8]{figures/rome}}
	\caption{Experimental results on the total resolution for the Rome graphs.}
	\label{fig:to-res}
\end{figure*}

% ==================================================================
%\input{an-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=9]{figures/rome}}
	\hfil
	\subfloat[\label{fig:an-res-2}{Edge-length ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=10]{figures/rome}}

	\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=11]{figures/rome}}
	\hfil
	\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=12]{figures/rome}}
	\caption{Experimental results on the angular resolution for the Rome graphs.}
	\label{fig:an-res}
\end{figure*}

@@

\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the results for the angular resolution; see Fig.~\ref{fig:an-res}. Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution). A notable observation is that, for graphs with less than $30$ vertices the best results are achieved by the algorithm by Argyriou et al., while for the remaining graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}. For large graphs, the two algorithms tend to have the same performance. The restricted variant of our algorithm yields drawings with slightly worse angular resolution. The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
%
The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.% This observation suggests that, for most of the graphs of our experiment, the angular resolution is greater than the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results. The small differences result from the fact that the crossing resolution cannot be entirely neglected.

>>>>

In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing. How would one increase the crossing resolution of a given drawing? First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase. Of course, we cannot consider all possible positions for the vertex to be moved. Instead, we consider a small set of randomly generated ones. If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position. Note that that in general the vertex movement paradigm is an established technique and has successfully applied to graph-theoretic and geometric problems; see, e.g.,~\cite{DBLP:conf/sofsem/BlasiusRR17,DBLP:conf/dac/FiducciaM82,6771089,DBLP:conf/alenex/RadermacherRRW18,DBLP:conf/wsdm/UganderB13}.

@@

% ==================================================================
%\input{cr-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=1]{figures/rome_colored}}
	\hfil
	\subfloat[\label{fig:cr-res-2}{Edge-length ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=2]{figures/rome_colored}}

	\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=3]{figures/rome_colored}}
	\hfil
	\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=4]{figures/rome_colored}}
	\caption{Experimental results on the crossing resolution for the Rome graphs.}
	\label{fig:cr-res}
\end{figure*}

% ==================================================================
%\input{to-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:to-res-1}{Total resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=9]{figures/rome_colored}}
	\hfil
	\subfloat[\label{fig:to-res-2}{Edge-length ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=10]{figures/rome_colored}}

	\subfloat[\label{fig:to-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=11]{figures/rome_colored}}
	\hfil
	\subfloat[\label{fig:to-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=12]{figures/rome_colored}}
	\caption{Experimental results on the total resolution for the Rome graphs.}
	\label{fig:to-res}
\end{figure*}

% ==================================================================
%\input{an-res-rome.tex}
% ==================================================================
\begin{figure*}[t]
	\centering
	\subfloat[\label{fig:an-res-1}{Angular resolution vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=5]{figures/rome_colored}}
	\hfil
	\subfloat[\label{fig:an-res-2}{Edge-length ratio vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=6]{figures/rome_colored}}

	\subfloat[\label{fig:an-res-3}{No.~of crossings vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=7]{figures/rome_colored}}
	\hfil
	\subfloat[\label{fig:an-res-4}{No.~of iterations vs no.~of vertices}]{
	\includegraphics[scale=0.99,page=8]{figures/rome_colored}}
	\caption{Experimental results on the angular resolution for the Rome graphs.}
	\label{fig:an-res}
\end{figure*}

@@

\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the results for the angular resolution; see Fig.~\ref{fig:an-res}. Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution). Our unrestricted algorithm yields drawings with better angular resolution than the corresponding ones produced by our restricted variant, and by the algorithms by Argyriou et al.\ and by Huang et al.; see Fig.~\ref{fig:an-res-1}. For graphs with 90 or more vertices, however, our unrestricted algorithms and the algorithm by Argyriou et al.\ tend to produce drawings with the same angular resolution on average. The restricted variant of our algorithm yields drawings with slightly worse angular resolution. The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
%
The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.% This observation suggests that, for most of the graphs of our experiment, the angular resolution is greater than the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results. The small differences result from the fact that the crossing resolution cannot be entirely neglected.
